Seismic Time Series Analysis

### 2020-06-15 project.Rmd
separate and better Lockdown impact

```{r show_overlap_and_diff, include=FALSE}
show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1) {
  overlap_plot <-
    autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
      geom_line(aes(y=base, col="red"), show.legend = F) +
      labs(title=NULL, subtitle=paste0(base.desc, " vs ", compare.desc))
  diff <- compare-base

>>>>

```{r show_overlap_and_diff, include=FALSE}
show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1, base.lab="red", compare.lab="black", overlap.legend=F) {
  overlap_plot <-
    autoplot(compare, xlab=xlab, ylab=overlap.ylab, series=compare.lab, show.legend = overlap.legend) +
      geom_line(aes(y=base, col=base.lab), show.legend = overlap.legend) +
      labs(title=NULL, subtitle=paste0(base.desc, " vs ", compare.desc))
  diff <- compare-base

### 2020-06-15 project.after.Rmd
separate and better Lockdown impact

plots <- show_overlap_and_diff(
  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
  base.desc = "test set during lockdown",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  overlap.ylab = data.lab,
  diff.title = NULL,
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))

>>>>

plots <- show_overlap_and_diff(
  final.testing, window(plotting.forecast$mean, start=c(71,1), end=c(78,168)), xlab=seasonality.xlab,
  base.desc = "test set before lockdown",
  base.lab = "Actual",
  compare.desc = paste0("forecast with ", plotting.forecast$method),
  compare.lab = "Forecast",
  overlap.legend = T,
  overlap.ylab = data.lab,
  diff.title = NULL,
  diff.ylab = "nm", diff.ylab.sec="%",
  diff.ylab.sec.accuracy = .01
)
print(plot_grid(plots[[1]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), plots[[2]] + geom_vline(aes(xintercept=lockdown.x), linetype=2), nrow = 2))
print(
  plots[[1]]
    + geom_vline(aes(xintercept=lockdown.x), linetype=2)
    + labs(title=paste0("Forecasts during lockdown with ", plotting.forecast$method), subtitle=NULL)
)

### 2020-06-15 slides.tex
separate and better Lockdown impact

\subsection{Seasonality}
\begin{frame}{\subsecname~- Visual inspection}
	\centering Plot review\\
	$\Downarrow$\\ %hints
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\caption{Zoom from January $6^{th}$ to February $3^{rd}$ 2019}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Periodgram}
	\centering Fourier Transform with different frequencies\\
	$\Downarrow$\\ %reveals
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/periodgram-1"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/periodgram-1"}
		\end{subfigure}
	\end{figure}
\end{frame}

\subsection{Decomposition}
\begin{frame}{\subsecname~- Trend extraction}\centering
	Tested simple, Spencer's and Moving Average filters
	\begin{figure}
		\includegraphics[width=.55\linewidth]{"project.uccs_files/figure-latex/trend MA-2"}
	\end{figure}
	\large$\Downarrow$\\\normalsize\textbf{Moving Average}\\\small better seasonal features
\end{frame}

\begin{frame}{\subsecname~- Seasonality extraction}
	\centering MA filter for detrending\\+\\local mean method%\\
%	single \& multiple seasonalities extraction
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/extract both 24 and 168-2"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/extract both 24 and 168-2"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize\textbf{Only weekly seasonality}\\\small fitting models
\end{frame}

\subsection{Lockdown impact}
\begin{frame}{\subsecname~- Multiple seasonalities}\centering
	Seasonalities extracted together
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize No significant difference
\end{frame}

\begin{frame}{\subsecname~- Weekly seasonality}\centering
	Only weekly seasonality extracted
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\vspace{.3em}\normalsize Monday to Friday lower values\\
	\vspace{.3em}Overall flattening towards 0
\end{frame}

@@

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-37"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-69"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\\vspace{1em}accepted model $\Rightarrow$\hspace{-1.3em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 17.48\% & $\approx81\%$ & $\approx91\%$ \\
				$(3,1,1)(0,1,1)_{168}$ & 15.18\% & $\approx96\%$ & $\approx98\%$ \\
				$(4,1,2)(0,1,1)_{168}$ & 14.68\% & $\approx98\%$ & $\approx99\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

@@

\begin{frame}{\subsecname~- Data transformation}\centering
	Spikes in data make it heavily non-Normal\\$\Downarrow$\\
	Transform using Box-Cox with $\lambda\approx-1$\\ %\quad\Rightarrow\quad y'=(y-1)/y$\\
	\vspace{.5em}\small Safe since all our data is positive\\\vspace{1em}\large$\Downarrow$\vspace{.5em}\\
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-7"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-9"}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-21"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-13"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\~\\~\\\vspace{1em}accepted model $\Rightarrow$ \hspace{-1.5em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$(0,1,0)(0,1,1)_{168}$ & 17.25\% & $\approx99\%$ & $100\%$ \\
				$(0,1,1)(0,1,1)_{168}$ & 17.98\% & $\approx98\%$ & $\approx99\%$ \\
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 16.14\% & $\approx73\%$ & $\approx89\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

\subsection{Lockdown impact}

\begin{frame}{\subsecname~- Forecasts assessment}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
		\end{subfigure}
	\end{figure}
	\begin{figure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
		\end{subfigure}
		\begin{subfigure}{.45\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models performance}
	\centering Different stations behaviour\\before lockdown $\rightarrow$ during lockdown\\~\\
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.3\linewidth}\centering
		\textbf{Uccle, Brussels}~~~
		\column{.15\linewidth}\centering
		\column{.35\linewidth}\centering
		\textbf{Membach}
		\column{.08\linewidth}
	\end{columns}
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.15\linewidth}\raggedleft
		${16\%} \rightarrow$\\
		${73\%} \rightarrow$\\
		${89\%} \rightarrow$\\
		\column{.12\linewidth}\raggedright
		$46\%$\\
		$39\%$\\
		$63\%$\\
		\column{.21\linewidth}\centering
		\textbf{$<$ MAPE $>$}\\\textbf{$<$ 80\% CI $>$}\\\textbf{$<$ 95\% CI $>$}\\
		\column{.17\linewidth}\raggedleft
		$17\% \rightarrow$\\
		$81\% \rightarrow$\\
		$91\% \rightarrow$\\
		\column{.15\linewidth}\raggedright
		${14\%}$\\
		${82\%}$\\
		${96\%}$\\
		\column{.08\linewidth}
	\end{columns}~\\
	\begin{columns}[T]
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Performance degradation
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Slight performance improvement\\
		\textit{lower peaks}
	\end{columns}
\end{frame}

>>>>

\subsection{Seasonality}
\begin{frame}{\subsecname~- Visual inspection}
	\centering Plot review\\
	$\Downarrow$\\ %hints
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
		\end{subfigure}
		\caption{Zoom from January $6^{th}$ to February $3^{rd}$ 2019}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Periodgram}
	\centering Fourier Transform with different frequencies\\
	$\Downarrow$\\ %reveals
	Daily and weekly seasonalities\\~
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/periodgram-1"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/periodgram-1"}
		\end{subfigure}
	\end{figure}
\end{frame}

\subsection{Decomposition}
\begin{frame}{\subsecname~- Trend extraction}\centering
	Tested simple, Spencer's and Moving Average filters
	\begin{figure}
		\includegraphics[width=.6\linewidth]{"project.uccs_files/figure-latex/trend MA-2"}
	\end{figure}
	\large$\Downarrow$\\\normalsize\textbf{Moving Average}\\\small better seasonal features
\end{frame}

\begin{frame}{\subsecname~- Seasonality extraction}
	\centering MA filter for detrending\\+\\local mean method%\\
%	single \& multiple seasonalities extraction
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/extract both 24 and 168-2"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/extract both 24 and 168-2"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize\textbf{Only weekly seasonality}\\\small fitting models
\end{frame}

@@

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-37"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/arima-69"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\\vspace{1em}accepted model $\Rightarrow$\hspace{-1.3em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 17.48\% & $\approx81\%$ & $\approx91\%$ \\
				$(3,1,1)(0,1,1)_{168}$ & 15.18\% & $\approx96\%$ & $\approx98\%$ \\
				$(4,1,2)(0,1,1)_{168}$ & 14.68\% & $\approx98\%$ & $\approx99\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

@@

\begin{frame}{\subsecname~- Data transformation}\centering
	Spikes in data make it heavily non-Normal\\$\Downarrow$\\
	Transform using Box-Cox with $\lambda\approx-1$\\ %\quad\Rightarrow\quad y'=(y-1)/y$\\
	\vspace{.5em}\small Safe since all our data is positive\\\vspace{1em}\large$\Downarrow$\vspace{.5em}\\
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-7"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/training data transformation plots-9"}
		\end{subfigure}
	\end{figure}
\end{frame}

\begin{frame}{\subsecname~- Models validation}
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-21"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/arima-13"}
		\end{subfigure}
	\end{figure}
	\begin{columns}[T]
		\column{.28\linewidth}\raggedleft~\\~\\~\\~\\\vspace{1em}accepted model $\Rightarrow$ \hspace{-1.5em}
		\column{.72\linewidth}
		\begin{table}
			\begin{tabular}{c | c c c}
				%			\toprule
				\textbf{ARIMA} & MAPE   & 80\% CI & 95\% CI \\
				\midrule
				$(0,1,0)(0,1,1)_{168}$ & 17.25\% & $\approx99\%$ & $100\%$ \\
				$(0,1,1)(0,1,1)_{168}$ & 17.98\% & $\approx98\%$ & $\approx99\%$ \\
				$\mathbf{(2,0,0)(0,1,1)_{168}}$ & 16.14\% & $\approx73\%$ & $\approx89\%$ \\
				%			\bottomrule
			\end{tabular}
		\end{table}
		\column{.05\linewidth}
	\end{columns}
\end{frame}

%------------------------------------------------
\section{Lockdown impact}
%------------------------------------------------

\subsection{Seasonalities}

\begin{frame}{Multiple seasonalities}\centering
	Seasonalities extracted together
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Uccle, Brussels}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
			\caption{Membach}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\normalsize No significant difference
\end{frame}

\begin{frame}{Weekly seasonality}\centering
	Only weekly seasonality extracted
	\begin{figure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-1"}
		\end{subfigure}
		\begin{subfigure}{.49\linewidth}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/seasonplot-mean-sd-2"}
		\end{subfigure}
	\end{figure}
	\large$\Downarrow$\\
	\vspace{.3em}\normalsize Monday to Friday lower values\\
	\vspace{.3em}Overall flattening towards 0
\end{frame}

\subsection{Forecasts}

\begin{frame}{Forecasts assessment}
	\begin{columns}[c]
		\column{.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
		\end{figure}
		\begin{figure}
			\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-10"}
		\end{figure}
		\column{.5\linewidth}\centering
		\textbf{Uccle, Brussels}\\
		lockdown $\rightarrow$ during lockdown\\~\\
		\begin{columns}[T]
			\column{.2\linewidth}
			\column{.33\linewidth}\raggedleft
			\textbf{MAPE}\\\textbf{80\% CI}\\\textbf{95\% CI}\\
			\column{.2\linewidth}\raggedleft
			${16\%} \rightarrow$\\
			${73\%} \rightarrow$\\
			${89\%} \rightarrow$\\
			\column{.1\linewidth}\raggedright
			$46\%$\\
			$39\%$\\
			$63\%$\\
			\column{.27\linewidth}
		\end{columns}~\\
		$\Downarrow$\\
		Performance degradation
	\end{columns}
\end{frame}

\begin{frame}{Forecasts assessment}
	\begin{columns}[c]
		\column{.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
		\end{figure}
		\begin{figure}
			\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-10"}
		\end{figure}
		\column{.5\linewidth}\centering
		\textbf{Membach}\\
		lockdown $\rightarrow$ during lockdown\\~\\
		\begin{columns}[T]
		\column{.2\linewidth}
			\column{.33\linewidth}\raggedleft
			\textbf{MAPE}\\\textbf{80\% CI}\\\textbf{95\% CI}\\
			\column{.2\linewidth}\raggedleft
			$17\% \rightarrow$\\
			$81\% \rightarrow$\\
			$91\% \rightarrow$\\
			\column{.1\linewidth}\raggedright
			${14\%}$\\
			${82\%}$\\
			${96\%}$\\
			\column{.27\linewidth}
		\end{columns}~\\
		$\Downarrow$\\
		Slight performance improvement\\
		\textit{lower peaks}
	\end{columns}
\end{frame}

\begin{frame}<0>[noframenumbering]{Models performance}
	\centering Different stations behaviour\\before lockdown $\rightarrow$ during lockdown\\~\\
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.3\linewidth}\centering
		\textbf{Uccle, Brussels}~~~
		\column{.15\linewidth}\centering
		\column{.35\linewidth}\centering
		\textbf{Membach}
		\column{.08\linewidth}
	\end{columns}
	\begin{columns}[T]
		\column{.12\linewidth}
		\column{.15\linewidth}\raggedleft
		${16\%} \rightarrow$\\
		${73\%} \rightarrow$\\
		${89\%} \rightarrow$\\
		\column{.12\linewidth}\raggedright
		$46\%$\\
		$39\%$\\
		$63\%$\\
		\column{.21\linewidth}\centering
		\textbf{$<$ MAPE $>$}\\\textbf{$<$ 80\% CI $>$}\\\textbf{$<$ 95\% CI $>$}\\
		\column{.17\linewidth}\raggedleft
		$17\% \rightarrow$\\
		$81\% \rightarrow$\\
		$91\% \rightarrow$\\
		\column{.15\linewidth}\raggedright
		${14\%}$\\
		${82\%}$\\
		${96\%}$\\
		\column{.08\linewidth}
	\end{columns}~\\
	\begin{columns}[T]
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Performance degradation
		\column{.5\linewidth}\centering
		$\Downarrow$\\
		Slight performance improvement\\
		\textit{lower peaks}
	\end{columns}
\end{frame}
